---
title: "Assignment 3"
author: "GP SINGH"
date: "February 11, 2016"
output: html_document
---

```{r}

library(RCurl)

```









```{r}

tb <- getURL("https://raw.githubusercontent.com/gpsingh12/IS-607-Assignment-3/master/tb.csv")
tb <- read.csv(text = tb, header = FALSE)
```

Assigning the names to the columns of dataset tb


```{r}
names(tb)<-c("country", "year", "sex", "child", "adult", "elderly")

head(tb)

```







As we looked at first few rows of tb, there were values that were NA giving us -1
in the columns. We need to replace these values for mathematical operations.


```{r}
tb <- replace(tb, tb == -1, 0)
head(tb)
```


creating new column by adding child, adult and elderly as cases


```{r}
tb <- transform(tb, case = child + adult + elderly)


# remove the columns  child, adult and elderly

tb$child<-NULL
tb$adult<-NULL
tb$elderly<-NULL

head(tb)
```




The the dataset tb contains both cases for male and female.  We create two datasets
with male and female.

```{r}
male_data <- subset(tb, tb$sex == "male")
female_data<-subset(tb, tb$sex == "female")

head(male_data)
head(female_data)
```



Add another column from female_data to male_date to add both the cases

```{r}
t1 <- transform(male_data, case1 = female_data$case)
head(t1)
```


here case represent male cases and case1 represent female cases


```{r}
# i will use matrix operations to add the columns  
t2<-transform(t1, cases = case+case1)


#remove columns sex, case and case1
t2$sex<-NULL
t2$case<-NULL
t2$case1<-NULL
# t2 contains the combined cases of male and female datasets
head(t2)
```


Reading data for population dataset


```{r}
p<- getURL("https://raw.githubusercontent.com/gpsingh12/IS-607-Assignment-3/master/population.csv")
population<- read.csv(text=p)
head(population)
```

Merge t2 and population datasets

```{r}

tb <- merge(t2, population, by=c("country", "year"))


#create a new column in tb by dividing cases by population

tb<-transform(tb,rate = (cases/population)*100000)


#  remove population and cases columns 

tb$population<-NULL
tb$cases<-NULL


#final result

head(tb)





```